CODECHECK: An open-science initiative for the independent execution of computations underlying research articles.

### 2021-01-07 codecheck-paper.tex
enough to send this out

\begin{document}
\title{CODECHECK: An open-science initiative to facilitate sharing of
  computer programs and results presented in scientific publications.}
\author[1,$\ast$]{Daniel N\"{u}st}
\author[2,$\ast$]{Stephen J Eglen}
\affil[1]{\url{https://orcid.org/0000-0002-0024-5046}}
\affil[2]{\url{https://orcid.org/0000-0001-8607-8025}}
\affil[$\ast$]{Contributed equally}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}}
\maketitle
\begin{abstract}
\ldots{}
\end{abstract}

@@

\begin{enumerate}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\item Should we make code and data sharing a prerequisite or leave that door open?
\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
\item Should we ask the people currently listed in the acknowledgements?
\item revisit \url{https://github.com/codecheckers/discussion}
\item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
\item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
\end{enumerate}

@@

Many areas of scientific research now use computation to either simulate
or analyse their data. These computations are increasingly complex, and
difficult to explain coherently in a paper \citep{marwick_how_2015}.
To complement the traditional route of sharing research by writing papers,
there is a growing trend/demand to share the underlying artifacts, notably 
code and datasets, so that others can inspect, reproduce or expand that work
(see Figure~\ref{fig:inverse}).
Some of the earliest proponents were Buckheit and Donoho
\cite{buckheit_wavelab_1995} who coined what has been called 
\emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{"An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures."}
%  Donoho (2010) has another paraphrasing of the same quote, which might work better:
% "an article about computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result."

So, assuming that researchers begin to share more artifacts, how might
they be examined to test that they do what they claim? For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share.
The implementation status of this varies considerably according to the
journal. At one of the spectrum, there are specialist journals that
have been created to accept ``data papers''
(e.g.~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the
other end of the spectrum, we still witness today the infamous
statement ``Data available upon reasonable request'' whereby authors,
while possible well-intentioned at the time of writing the article,
often cannot provide the data when readers ask for
it.  (ANDREW H: any references to sum up poor/evolving state of data
sharing; Vines et al 2014).

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? While both data and software
together are required to validate outcomes of a computational analysis,
the do differ in the respect that the former can be seen as static/inert,
while the latter requires an environment and interaction. This makes 
software harder to share.
Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in Open Science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/execute
the code, this is worth documenting as early as possible and that the barrier
to evaluating non-text parts of research can be greatly reduced with a 
focused approach.
Just like peer review can provide a ``baseline reasurance'', i.e., a paper
has at least ben checked by someone with an understanding of the topic
\cite{fyfe_mission_2019}, we think the same baseline should be provided
regarding the workflow underlying a paper.
With this in mind, we have developed a set of PRINCIPLES and an example 
workflow that provides what he hope is a pragmatic way of checking that 
code works.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research.  The left
  half of the diagram shows the diverse range of materials used
  internally within a laboratory.  These materials are often then
  condensed into a paper for sharing with the outside world via the
  research paper, a static PDF document.  Working backwards from the
  PDF to the underlying materials is impossible. This probihibits reuse
  and is not only non-transparent for a specific paper, but also 
  ineffective for science as a whole. By sharing the
  materials on the left, others outside the lab are able to reproduce
  or build on this work.}
  \label{fig:inverse}
\end{figure}

@@

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people: (1) the AUTHOR of a paper providing the code to be checked, (2)
the PUBLISHER of a journal intereseted in publishing the paper by the
AUTHOR and (3) the CODECHECKER who checks that the AUTHOR's code works.
The workflow that we have refined is documented in Figure~\ref{fig:workflow}.

@@

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
% emoji codes:
% book stack: 1f4da
% package: 1f4e6
% file drawer: 1f5c4
% disk: 1f4be
% document: 1f4c4
% phd: 1f393
% detective: u1f575
% laptop: 1f4bb
% folder: 1f4c1 1f4c2
% receipt: 1f9fe
  \centering
      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
  \caption{The CODECHECK example process implementation. Codechecker act as a detectives:
  The investigate and record, but do not fix issues.}
  \label{fig:worfklow}
\end{figure}

@@

\textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
analogous to the publisher finding one or more reviewers to evaluating
the paper. However, a cruical difference is that we suggest the CODECHECKER
and the author can talk freely to each other directly (see next).

\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
the author, and checks if some/all of the results from the paper can be
reproduced. (See later for exact nature of ``reproduced''). If there are
any problems running the code, the CODECHECKER asks the author for help
to resolve the problems. The burden to provide reproducible material lies with
the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code does not work (\emph{what do in this
case}). As part of this process, the CODECHECKER could work entirely
locally on their own compute resource, or in the cloud, e.g.~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
are more tailored for scientific publications while others offer commercial
options to, e.g., publishers (cf. \cite{konkol_publishing_2020}).
Using such cloud-based 
infrastructure allows for the CODECHECKER and author to collaboratively improve
the code and enforces a much more complete definition of the computing environment,
but, unless infrastructure is provided by, e.g., by the publisher, requires code
and data to be published openly online.
Note that the taks of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} destinguish between verification and
validation in the context of Mathematics, and in the same vein a CODECHECK
ensures a verification\footnote{Note this is a different use as the \emph{Verification Reports} article type in the journal \emph{Cortex}
\cite{chambers_verification_2020}},
i.e., checking that the code correctly produces the 
output it claims to create, but not a validation, i.e., does the code implement
the right algorithm to solve the problem that is researched, 
of the computational results.
Nevertheless, the simple attempt to reproduce may highlight shortcomings in
a submission with respect to transparency as it may be required by a journal's
policy and help to resolve misunderstandings and discrepancies
(cf.~\cite{christian_journal_2020}).

@@

\textbf{Step 5:} the certificate as well as auxiliary files created during the check,
e.g., a specifiction of a computing environment, helper scripts, and the code and data,
if available under open licenses, get deposited in an open archive (currently Zenodo);
the certificate is given to the PUBLISHER.

\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
supplementary records for a paper or provide it to scientific reviewers; the
PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
activity appropriately in scholarly publication databases, such as ORCID.
The PULISHER also ensures proper publication metadata, e.g., links from the 
certificate repository to the published paper or the original code repository.
A badge or other visual aid may be applied to highlight the fact of a completed
CODECHECK and provide a incentive and reward for the, today often considered
``additional'' efforts of providing readily computationally reproducible workflows,
despite the simplification into a binary value and the possible confusion of
the meaning, especially across disciplines and methods.

@@

@@

A pre-review CODECHECK allows editors to use it as a filter to only send 
a submission out to peer review if it passes the check. A less strict
pre-review check would simply be included in the submission package provided
to the reviewers, so they can include the CODECHECK's results in their 
evaluation of the work.
A CODECHECK may also be conducted in parallel to the scientific review. This
puts less burden on the turnaround time for the check yet only makes the 
outcomes available for the final consideration of the handling editor.
The check could also be tasked on demand, e.g., only if one of the reviewers
suggests/requests one because previous screenings might not have caught the 
fact that a computational workflow is part of the submission. However, 
soliciting such a "specialist review" is much more undesirable than having
a regular CODECHECK, without any special treatment for some submissions.
A CODECHECK could then be awared equal weight as the other reviews.
A post acceptance CODECHECK puts the least requirements on the author and may
simply lead to an award of excellence on top of the acceptance of the
submission. This is the least strict solution while still properly 
acknowleding the check results, because the CODECHECK is to be completed 
before the publication of the final paper.
The GIScience group of checks (see \nameref{register}) falls into this
category: the AGILE conference highlights articles where the reproducibility
was successfully reviewed with a badge on the volume and article landing
pages.

@@

The options for the person who can conduct a CODECHECK require a not quite
simple matching of a combination of skills and availability. Ideally, the
codechecker has a matching code \emph{and} domain expertise, for example to
check a workflow based on Python code analysing MRT brain imagery. However,
a well-documented workflow should be executable by any person with a basic
understanding of computers. Naturally, the more prerequisite knowledge the
codechecker has, the less time is spend to understand the goals and 
the mechanics of an analysis. From our experiences, the priority should be
given to matching technical expertise first, as the a lack of knowledge in
setting  up a computing environment with a particular language or tool is 
much more discouraging and the assessment of the outcome, e.g., comparing
created figures with the original, is largely possible without 
understanding.
The time and expertise alloted to the check
are the main drivers for the depth of checking, though in general, we 
expect a CODECHECK not to evaluate performance, e.g., does a racing car
have the best apex speed, but to assess operation free of failure, i.e., 
is it a car with 4 wheels and at least one door, and does it move forwards
when accelerating. This depth can be seen independent from the power of
the check (see above), as long as it is communicated clearly to all
stakeholders, most importantly authors and readers.

The concrete people to provide these expertises can range from researchers,
possibly drawn from the regular pool of volunteer reviewers or from a 
special group, to editorial staff of a publisher. If the regular reviewers
are relied upon, the editor's task to find suitable ones becomes more
tedious; extra information such as software familiarity would have to be
modelled in review platforms. Keeping diverse specialists as staff to
realise a codechking process is unlikely to be feasible, especially for 
small or independent publishers.

We consider a single codechecker to be enough, unlike peer review where
multiple referees are common. Because of the technical focus, the 
determination if the concrete outputs of author and codechecker, who can
collaborate to solve any issues, should be a very factual process and not
require much debate or opinion. Depending on the power of the CODECHECK
(see \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
second opinion might be warranted in rare cases.

We see a great opportunity to especially involve early-career researchers
(ECRs) as codecheckers. They arguably have a high interest in learning
about the latest tools and technologies as they are building up their own
expertise and specialisation. The possibility to gain insights into latest
science for ECRs and at the same time increase the appreciation of
reproduction efforts is a sweet spot that CODECHECK shares
with \emph{ReScience X}, a journal devoted to reproduction and replication 
experiments \cite{roesch_new_2020}.
In fact, ECRs are often much more familiar with latest technology and are 
likely to be early adoptors as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
submitted to leading Neuroscience journals by ECRs
\cite{chambers_registered_2019}.}
They are also introduced to peer review and 
may transition into the role of scientific reviewer over time.
Not to solve the general problems of peer review, which is largely an
unsupervised process, i.e., rarely are reviewers taught how to do it, but
a junior codechecker may be supported by a senior codechecker, too.
We see a high potential in setting up the new processes for CODECHECKING with
a clear committment on openness and transparency, independent of the
respective current peer review process (see \nameref{who-knows-who}).

Publisher staff represents the most controlled yet also not quickly scalable
option. Hiring the required technical and or domain expertise puts a financial
burden on the publisher, of course, but such a committment also shows the 
value given to publishing reproducibility and would positively impact the 
submission experience, because of more controllability, and external 
perception of the journal or conference. A staff codechecker reduces the 
challenges when it comes to open communication and awarding credit, as
in-house personnel may be more readily be given access to non-anonymised
submission material and receives a proper salary instead of public credit
as part of the scientific community.

In contrast, for researchers it can be very important to be publicly credited
with their activity as a reviewer (see \nameref{who-knows-who}). A regular
review may be listed in public databases (e.g., ORCID
\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
\footnote{\href{https://publons.com/}{https://publons.com/}} or 
ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
A codechecker should be listed just the same. The number of scientific reviews
is of course a very coarse indicator for the community contribution by an 
individual, yet is is unclear if a regular reviewer, who in addition
conducts a CODECHECK to the scientific review, should be credited twice.

@@

The topic of blinding is broadly discussed, not the least in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stand on this discussion, the general motivation 
for more  transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can also serve to protect individuals 
\ref{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
be acceptable to require revealing of the names of tenured researchers.
The possible negative effects of a non-anonymous review are greatly reduced
if a CODECHECK is not relevant for the decision to accept or reject, but that
is of course not desirable. Instead, we argue that a CODECHECK is technical
process that should in any case be fixable and not a question of opinion or
faulty approach. It is possible for a nonsense workflow to receive a 
CODECHECK certificate.
Furthermore, the level of documentation that is needed for third parties
to reproduce a workflow is extremely hard to get right. Too often, this 
uncertainty leads to researchers giving up and not documenting at all.

The technical nature of the check and the challenge of good enough 
documentation is why we see great benefits in a effective bidirectional
means of communication between author and codechecker. Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the docs  or the code.
Instead of lenghtily describing a small change, the codechecker may simply
provide a fix within the code.
Instead of hiding useful and  instructive files created during the codecheck
(e.g., a machine-readable computing environment specification), the author 
and readers can profit from the software-related expertise of the
codechecker.
While such a communication may be facilitated in an anonymous way by the 
publisher, it most likely only helps to protect the identity of the 
codechecker, because, especially when of a certain size, not trivial, or 
if very specialised, it is close to impossible to fully anonymise a 
software analysis stack.
Therefore, the most effective means with the most desirable result for
the stakeholders is a non-anonymous and collaborative CODECHECK.
The possible contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit is given within the research software development community.
Nevertheless, depending on other needs represented by other dimensions,
the aspect of `Who knows who?' may be the most flexible one.

@@

\textbf{2. Communication between humans is key.} Some code may just work
without any interaction (e.g.~Manuel's certificate TODO add cert number); 
however, more often
there are hidden dependencies that need to be found, or revised software
installations. Allowing the codechecker and human to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would slow things down and
reduce the chances for community building as well as the lessons learned
for both author and codecheckers.

>>>>

\begin{document}
\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles.}
\author[1,$\ast$]{Daniel N\"{u}st}
\author[2,$\ast$]{Stephen J Eglen}
\affil[1]{\url{https://orcid.org/0000-0002-0024-5046}}
\affil[2]{\url{https://orcid.org/0000-0001-8607-8025}}
\affil[$\ast$]{Contributed equally}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}}
\maketitle
\begin{abstract}
  This is a draft paper that we hope to submit to F1000R (and perhaps
  as a preprint before) in the coming weeks.
  \ldots{}
\end{abstract}

@@

\begin{enumerate}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Can we introduce "research compendium" as a term?  (yes).
\item Should we make code and data sharing a prerequisite or leave
  that door open? (make it the norm)
\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
\item Check with people  in the acknowledgements
%\item revisit \url{https://github.com/codecheckers/discussion}
% \item \url{https://hdsr.mitpress.mit.edu/pub/f0obb31j/release/1} and other articles form the HDSR special issue
% \item \url{https://av.tib.eu/media/42484} > "Code will make systematic mistakes, not random mistakes", "code is more reliable (and auditable) than a human"
\end{enumerate}

@@

Many areas of scientific research now use computation to either simulate
or analyse their data. These computations are increasingly complex, and
difficult to explain coherently in a paper \citep{marwick_how_2015}.
To complement the traditional route of sharing research by writing papers,
there is a growing trend/demand to share the underlying artefacts, notably 
code and datasets, so that others can inspect, reproduce or expand that work
(see Figure~\ref{fig:inverse}).
Some of the earliest proponents were Buckheit and Donoho
\cite{buckheit_wavelab_1995} who coined what has been called 
\emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{"An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures."}
%  Donoho (2010) has another paraphrasing of the same quote, which might work better:
% "an article about computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result."

So, assuming that researchers begin to share more artefacts, how might
they be examined to test that they do what they claim? For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share.
The implementation status of this varies considerably according to the
journal. At one of the spectrum, there are specialist journals that
have been created to accept ``data papers''
(e.g.~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the
other end of the spectrum, we still witness today the infamous
statement ``Data available upon reasonable request''. Authors,
while possible well-intentioned at the time of writing the article,
often cannot provide the data when readers ask for
it as data disappears over time \cite{Vines2014-hf}.

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? While both data and software
together are required to validate outcomes of a computational analysis,
they do differ in  that data can be seen as static/inert,
while code requires an environment and interaction. This makes 
software harder to share.
Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in Open Science principles, is that
wherever possible sharing code is good for the community, as outlined
in 2010 \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/execute
the code, this is worth documenting as early as possible and that the barrier
to evaluating non-text parts of research can be greatly reduced with a 
focused approach.
Just like peer review can provide a ``baseline reassurance'', i.e., a paper
has at least been checked by someone with an understanding of the topic
\cite{fyfe_mission_2019}, we think the same baseline should be provided
regarding the workflow underlying a paper.
With this in mind, we have developed a set of PRINCIPLES and an example 
workflow that provides what he hope is a pragmatic way of checking that 
code works.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research.  The left
  half of the diagram shows the diverse range of materials used
  within a laboratory.  These materials are often then
  condensed for sharing with the outside world via the
  research paper, a static PDF document.  Working backwards from the
  PDF to the underlying materials is impossible. This prohibits reuse
  and is not only non-transparent for a specific paper, but also 
  ineffective for science as a whole. By sharing the
  materials on the left, others outside the lab can reproduce
  or build on this work.}
  \label{fig:inverse}
\end{figure}

@@

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people: (1) the AUTHOR of a paper providing the code to be checked, (2)
the PUBLISHER of a journal interested in publishing the paper by the
AUTHOR and (3) the CODECHECKER who checks that the AUTHOR's code works.
The workflow that we have refined is documented in Figure~\ref{fig:workflow}.

@@

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
% emoji codes:
% book stack: 1f4da
% package: 1f4e6
% file drawer: 1f5c4
% disk: 1f4be
% document: 1f4c4
% phd: 1f393
% detective: u1f575
% laptop: 1f4bb
% folder: 1f4c1 1f4c2
% receipt: 1f9fe
  \centering
      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
  \caption{The CODECHECK example process implementation. Codechecker act as a detectives:
  They investigate and record, but do not fix issues.}
  \label{fig:workflow}
\end{figure}

@@

\textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
analogous to the publisher finding one or more reviewers to evaluating
the paper. However, a crucial difference is that we suggest the CODECHECKER
and the author can talk freely to each other directly (see next).

\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
the author, and checks if some/all of the results from the paper can be
reproduced. (See later for exact nature of ``reproduced''). If there are
any problems running the code, the CODECHECKER asks the AUTHOR for help
to resolve the problems. The burden to provide reproducible material lies with
the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code does not work (\emph{TODO NOT
  YET HANDLED}). As part of this process, the CODECHECKER could work entirely
locally on their own compute resource, or in the cloud, e.g.~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
are more tailored for scientific publications while others offer commercial
options to, e.g., publishers (cf. \cite{konkol_publishing_2020}).
Using such cloud-based 
infrastructure allows for the CODECHECKER and author to collaboratively improve
the code and enforces a much more complete definition of the computing environment,
but, unless secure infrastructure is provided by, e.g., by the publisher, requires code
and data to be published openly online.
Note that the task of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} distinguish between verification and
validation in the context of Mathematics, and in the same vein a CODECHECK
ensures a verification\footnote{Note this is a different use as the \emph{Verification Reports} article type in the journal \emph{Cortex}
\cite{chambers_verification_2020}},
i.e., checking that the code correctly produces the 
output it claims to create, but not a validation, i.e., does the code implement
the right algorithm to solve the problem that is researched, 
of the computational results.
Nevertheless, the simple attempt to reproduce may highlight shortcomings in
a submission with respect to transparency as it may be required by a journal's
policy and help to resolve misunderstandings and discrepancies
(cf.~\cite{christian_journal_2020}).

@@

\textbf{Step 5:} the certificate as well as auxiliary files created
during the check, e.g., a specification of a computing environment,
helper scripts, and the code and data, if available under open
licenses, get deposited in an open archive (currently Zenodo); the
certificate is given to the PUBLISHER.

\textbf{Step 6:} the PUBLISHER can then use the certificate as part of their
supplementary records for a paper or provide it to scientific reviewers; the
PUBLISHER also gives credit to the CODECHECKER for their work by depositing the
activity appropriately in scholarly publication databases, such as ORCID.
The PUBLISHER also ensures proper publication metadata, e.g., links from the 
certificate repository to the published paper or the original code repository.
A badge or other visual aid may be applied to highlight the fact of a completed
CODECHECK and provide a incentive and reward for the, today often considered
``additional'' efforts of providing readily computationally reproducible workflows,
despite the simplification into a binary value and the possible confusion of
the meaning, especially across disciplines and methods.

@@

\textbf{Daniel to fix by merging turnaround time into other
  dimensions, and labelling these as the four dimensions: When?
  Openness? Who? Importance?}

@@

A pre-review CODECHECK allows editors to use it as a filter to only send 
a submission out to peer review if it passes the check. A less strict
pre-review check would simply be included in the submission package provided
to the reviewers, so they can include the CODECHECK's results in their 
evaluation of the work.
A CODECHECK may also be conducted in parallel to the scientific review. This
puts less burden on the turnaround time for the check yet only makes the 
outcomes available for the final consideration of the handling editor.
The check could also be tasked on demand, e.g., only if one of the reviewers
suggests/requests one because previous screenings might not have caught the 
fact that a computational workflow is part of the submission. However, 
soliciting such a "specialist review" is much more undesirable than having
a regular CODECHECK, without any special treatment for some submissions.
A CODECHECK could then be awarded equal weight as the other reviews.
A post acceptance CODECHECK puts the least requirements on the author and may
simply lead to an award of excellence on top of the acceptance of the
submission. This is the least strict solution while still properly 
acknowledging the check results, because the CODECHECK is to be completed 
before the publication of the final paper.
The GIScience group of checks (see \nameref{register}) falls into this
category: the AGILE conference highlights articles where the reproducibility
was successfully reviewed with a badge on the volume and article landing
pages.

@@

The options for the person who can conduct a CODECHECK require a not quite
simple matching of a combination of skills and availability. Ideally, the
codechecker has a matching code \emph{and} domain expertise, for example to
check a workflow based on Python code analysing neuroimaging datasets. However,
a well-documented workflow should be executable by any person with a basic
understanding of computers. Naturally, the more prerequisite knowledge the
codechecker has, the less time is needed to understand the goals and 
the mechanics of an analysis. From our experiences, the priority should be
given to matching technical expertise first, as the a lack of knowledge in
setting  up a computing environment with a particular language or tool is 
much more discouraging and the assessment of the outcome, e.g., comparing
created figures with the original, is largely possible without 
understanding.
The time and expertise allotted to the check
are the main drivers for the depth of checking, though in general, we 
expect a CODECHECK not to evaluate performance, e.g., does a racing car
have the best apex speed, but to assess operation free of failure, i.e., 
is it a car with 4 wheels and at least one door, and does it move forwards
when accelerating. This depth can be seen independent from the power of
the check (see above), as long as it is communicated clearly to all
stakeholders, most importantly authors and readers.

The concrete people to provide these expertises can range from researchers,
possibly drawn from the regular pool of volunteer reviewers or from a 
special group, to editorial staff of a publisher. If the regular reviewers
are relied upon, the editor's task to find suitable ones becomes more
tedious; extra information such as software familiarity would have to be
modelled in review platforms. Keeping diverse specialists as staff to
realise a codechecking process is unlikely to be feasible, especially for 
small or independent publishers.

We consider a single codechecker to be enough, unlike peer review where
multiple referees are common. Because of the technical focus, the 
determination if the concrete outputs of author and codechecker, who can
collaborate to solve any issues, should be a very factual process and not
require much debate or opinion. Depending on the power of the CODECHECK
(see TODO: check ref \ref{when-to-do-a-codecheck}), a mediator, e.g., the editor, or a
second opinion might be warranted in rare cases.

We see a great opportunity to especially involve early-career researchers
(ECRs) as codecheckers. They arguably have a high interest in learning
about the latest tools and technologies as they are building up their own
expertise and specialisation. The possibility to gain insights into latest
science for ECRs and at the same time increase the appreciation of
reproduction efforts is a sweet spot that CODECHECK shares
with \emph{ReScience X}, a journal devoted to reproduction and replication 
experiments \cite{roesch_new_2020}.
In fact, ECRs are often much more familiar with latest technology and are 
likely to be early adopters as creators, i.e., be authors of CODECHECK-ready manuscripts, too\footnote{Cf. the 77\% of 141 registered reports who were 
submitted to leading Neuroscience journals by ECRs
\cite{chambers_registered_2019}.}
They are also introduced to peer review and 
may transition into the role of scientific reviewer over time.
Not to solve the general problems of peer review, which is largely an
unsupervised process, i.e., rarely are reviewers taught how to do it, but
a junior codechecker may be supported by a senior codechecker, too.
We see a high potential in setting up the new processes for CODECHECKING with
a clear commitment on openness and transparency, independent of the
respective current peer review process (see \nameref{who-knows-who}).

Publisher staff represents the most controlled but also resource-intensive
option. Hiring the required technical and or domain expertise puts a financial
burden on the publisher, of course, but such a commitment also shows the 
value given to publishing reproducibility and would positively impact the 
submission experience, because of more controllability, and external 
perception of the journal or conference. A staff codechecker reduces the 
challenges when it comes to open communication and awarding credit, as
in-house personnel may be more readily be given access to non-anonymised
submission material and receives a proper salary instead of public credit
as part of the scientific community.

In contrast, for researchers it can be very important to be publicly credited
with their activity as a reviewer (see TODO CHECK REF \nameref{who-knows-who}). A regular
review may be listed in public databases (e.g., ORCID
\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}} or commercial offerings such as Publons
\footnote{\href{https://publons.com/}{https://publons.com/}} or 
ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}}).
A codechecker should be listed just the same. The number of scientific reviews
is of course a very coarse indicator for the community contribution by an 
individual, yet it is unclear if a regular reviewer, who in addition
conducts a CODECHECK to the scientific review, should be credited twice.

@@

The topic of blinding is broadly discussed, not the least in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stance on this discussion, the general motivation 
for more  transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can also serve to protect individuals 
\cite{tennant_limitations_2020}, e.g.,  junior scientists, whereas it might 
be acceptable to require revealing of the names of tenured researchers.
The possible negative effects of a non-anonymous review are greatly reduced
if a CODECHECK is not relevant for the decision to accept or reject, but that
is of course not desirable. Instead, we argue that a CODECHECK is technical
process that should in any case be fixable and not a question of opinion or
faulty approach. It is possible for a nonsense workflow to receive a 
CODECHECK certificate.
Furthermore, the level of documentation that is needed for third parties
to reproduce a workflow is extremely hard to get right. Too often, this 
uncertainty leads to researchers giving up and not documenting at all.

The technical nature of the check and the challenge of good enough 
documentation is why we see great benefits in a effective bidirectional
means of communication between author and codechecker. Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the docs  or the code.
Instead of lengthily describing a small change, the codechecker may simply
provide a fix within the code.
Instead of hiding useful and  instructive files created during the codecheck
(e.g., a machine-readable computing environment specification), the author 
and readers can profit from the software-related expertise of the
codechecker.
While such a communication may be facilitated in an anonymous way by the 
publisher, it most likely only helps to protect the identity of the 
codechecker, because, especially when of a certain size, not trivial, or 
if very specialised, it is close to impossible to fully anonymise a 
software analysis stack.
Therefore, the most effective means with the most desirable result for
the stakeholders is a non-anonymous and collaborative CODECHECK.
The possible contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit is given within the research software development community.
Nevertheless, depending on other needs represented by other dimensions,
the aspect of `Who knows who?' may be the most flexible one.

@@

\textbf{2. Communication between humans is key.} Some code may just
work without any interaction (e.g. CODECHECK certificate 2020-013
\cite{cert-2020-013}); however, more often there are hidden
dependencies that need to be found, or revised software
installations. Allowing the codechecker and human to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would slow things down and
reduce the chances for community building as well as the lessons
learned for both author and codecheckers.

\textbf{3. Credit is given to codecheckers.} The value of contributing a
CODECHECK is comparable to providing a scientific peer review. Individual
variations aside, it may also require a similar amount of time. Therefore,
the codechecker's activity should be publicly recorded and mentioned in the
published paper.
The public record can be realised by publishing the certificate in a 
citable form (i.e., with a DOI; this is done within the community codecheck
process), by listing codecheckers on the journal's website or, ideally, by
publishing the checks alongside peer review activities in public databases.
